---
title: "Data Extraction"
author: "Joseph Blubaugh"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: DejaVu Sans Mono
sansfont: DejaVu Sans Mono
fontsize: 11pt
geometry: margin=.5in
---

Since the data is not really big, rather than writing a full program to deal with the different formatting of the various sheets I instead did a little prep work on each sheet to make it a little easier to read into R. I created columns for County, Treatment, and Setup Day then I transposed the Week, Date, and Count data so that the week indicators was in a single column. This helped to flatten the data a bit.

```{r a, comment=NA, message=FALSE, warning=FALSE}
library(xlsx)
library(reshape2)
library(knitr)

dta = read.xlsx(file = "../FlyData.xlsx", 
                startRow = 1, endRow = 49, 
                colIndex = 1:15, 
                sheetName = "Anderson", 
                header = TRUE)

kable(head(dta))

```

In the R package **reshape2** there is a function called **melt** which can easily be used to collapse the data further into something more usable.

```{r b, comment=NA, message=FALSE, warning=FALSE}

dta = melt(data = dta, id.vars = c("County", "Treatment", "Setup", "Week", "Date"), 
           variable.name = "Measurement", 
           value.name = "Count")

kable(head(dta))
```

