---
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: Calibri Light
sansfont: Calibri Light
fontsize: 11pt
geometry: margin=1in
---

**Homework 09**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 659-700**  

```{r a, echo=FALSE, warning=FALSE, message=FALSE}
library(pander)
```


### 5.21

```{r b, comment=NA, warning=FALSE, message=FALSE}
(alcohol = matrix(c(0, .5, 1.5, 5, 7, 48, 38, 5, 1, 1), nrow = 5))

fisher.test(alcohol)
```

### 5.22

```{r c, comment=NA, results='asis', echo=FALSE}
dta = data.frame(
  alcohol = c(0, .5, 1.5, 4, 7),
  absent = c(17066, 14464, 788, 126, 37),
  present = c(48, 38, 5, 1, 1)
)

dta = data.frame(
  y = c(0, 0, 0, 0, 1, 1, 1, 1),
  x = c(10, 20, 30, 40, 60, 70, 80, 90)
)

pandoc.table(dta, caption = "malformation")
```

\newpage

a. $\beta$ is nearly 0 because there is a clear distinction between y = 0 and y = 1.

```{r d, comment=NA, echo=FALSE}

mdl = glm(y ~ x, family = binomial(), data = dta)
summary(mdl)

```

b. reported above

c. No because there is still a clear distinction in x when y = 0 or y = 1.

```{r e, comment=NA, echo=FALSE}
dta = data.frame(
  y = c(0, 0, 0, 0, 1, 1, 1, 1, 0, 1),
  x = c(10, 20, 30, 40, 60, 70, 80, 90, 50, 50)
)

mdl = glm(y ~ x, family = binomial(), data = dta)
summary(mdl)

```

d. Yes, this is correct because now there is overlap between the point where y = 0 or y = 1 conditional on x. In the other two examples there is not "gray" area so there is nothing to model essentially.

```{r f, comment=NA, echo=FALSE}
dta = data.frame(
  y = c(0, 0, 0, 0, 1, 1, 1, 1, 0, 1),
  x = c(10, 20, 30, 40, 60, 70, 80, 90, 50.1, 49.9)
)

mdl = glm(y ~ x, family = binomial(), data = dta)
summary(mdl)

```


### 5.23

```{r g, comment=NA, echo=FALSE, results='asis'}

dta = data.frame(
  penicillin = c("1/8", "1/8", "1/4", "1/4", "1/2", "1/2", "1", "1", "4", "4"),
  delay = rep(c("None", "1.5hrs"), 5),
  cured = c(0, 0, 3, 0, 6, 2, 5, 6, 2, 5),
  died = c(6, 5, 3, 6, 0, 4, 1, 0, 0, 0)
)

dta$penicillin = factor(dta$penicillin, levels = c("1/8", "1/4", "1/2", "1", "4"))
dta$delay = factor(dta$delay, levels = c("None", "1.5hrs"))

pandoc.table(dta)

```

a. The software reports highly insignificant coefficients for both the first and last penicillin doses. The coefficients are large, but the standard errors are very large and include 0

```{r h, comment=NA}
mdl = glm(cbind(dta$cured, dta$died) ~ penicillin + delay - 1, family = binomial(), 
          data = dta)
summary(mdl)
```

b. Based on a pvalue < .05 we would conclude that the odds ratios condtional on penicillin level are not equal. $3.94 > 3.84 = qchisq(.95, 1)$

```{r i, comment=NA, warning=FALSE, message=FALSE}
library(lawstat)
library(reshape2)

dta = melt(data = dta, id.vars = c("penicillin", "delay"), 
           variable.name = "status", 
           value.name = "freq")
dta = xtabs(freq ~ delay + status + penicillin, data = dta)
mantelhaen.test(dta)

```

c. The conditional odds ratios are not estimatable because every level of penicillin has at least one 0.

```{r j, comment=NA, warning=FALSE, message=FALSE}
cmh.test(dta)

```


### 5.24

Since we are dealing with several cells that have 0, we can pool the results over the levels and estimate that the odds ratio is 1.84.


### 6.1

a. $log(\frac{\pi_R}{\pi_D}) = log(\frac{\pi_R}{\pi_I}) - log(\frac{\pi_D}{\pi_I}) = (1 + .3x) - (3.3 - .2x) = -2.3 + .5x$
Holding income constant the odds of someone prefering Republican vs Demovcrat is exp(-2.3) = .091

b. $log(\frac{\pi_R}{\pi_D}) = 0 = -2.3 + .5x \rightarrow 2.3 = .5x \rightarrow x = 2.3/5 = .46$ The point at which the probability of preferring Republican is higher is when income reaches $46K.

c. $\pi_I = 1 / 1 + e^{1 + .3x} + e^{3.3 - .2x}$

\newpage

### 6.2

a. Between Invertebrates or Other the odds of a Aligator preferring Invertibrates over Other increases exp(-2.46) = .085 for unit increase in length

```{r k1, comment=NA, echo=FALSE}
gator = data.frame(
  length = c(1.24, 1.30, 1.30, 1.32, 1.32, 1.40, 1.42, 1.42, 1.45, 
             1.45, 1.47, 1.47, 1.50, 1.52, 1.55, 1.60, 1.63, 1.65, 
             1.65, 1.65, 1.65, 1.68, 1.70, 1.73, 1.78, 1.78, 1.78, 
             1.80, 1.80, 1.85, 1.88, 1.93, 1.98, 2.03, 2.03, 2.16, 
             2.26, 2.31, 2.31, 2.36, 2.36, 2.39, 2.41, 2.44, 2.46, 
             2.56, 2.67, 2.72, 2.79, 2.84, 3.25, 3.28, 3.33, 3.56, 
             3.58, 3.66, 3.68, 3.71, 3.89),
  pref = c("Invertebrate", "Invertebrate", "Invertebrate", "Fish", "Fish", "Fish", 
           "Invertebrate", "Fish", "Invertebrate", "Other", "Invertebrate", "Fish", 
           "Invertebrate", "Invertebrate", "Invertebrate", "Invertebrate", 
           "Invertebrate", "Other", "Invertebrate", "Fish", "Fish", "Fish", 
           "Invertebrate", "Other", "Invertebrate", "Invertebrate", "Other", 
           "Invertebrate", "Fish", "Fish", "Invertebrate", "Invertebrate", 
           "Invertebrate", "Fish", "Fish", "Fish", "Fish", "Fish", "Fish", "Fish", 
           "Fish", "Fish", "Fish", "Fish", "Fish", "Other", "Fish", "Invertebrate", 
           "Fish", "Fish", "Other", "Other", "Fish", "Fish", "Fish", "Fish", "Other", 
           "Fish", "Fish")
  )

gator$pref = relevel(gator$pref, ref = "Other")

```

```{r k2, comment=NA}

library(nnet)

summary(gator)

mdl = multinom(pref ~ length, data = gator); summary(mdl)


```

b. The probability that the gator will prefer fish is .762

```{r l, comment=NA}
predict(mdl, data.frame(length = 3.9), type = "probs")
```

c. $log(\frac{\pi_I}{\pi_O}) = 0 = 5.697 - 2.465x \rightarrow -5.697 = -2.465x \rightarrow x = 2.31$


### 6.3

```{r m, comment=NA, echo=FALSE}

gator = 
  data.frame(
    lake = c("H", "H", "H", "H", "H", "H", "H", "H", "H", "H", "O", "O", "O", 
             "O", "O", "O", "O", "O", "O", "O", "T", "T", "T", "T", "T", "T", 
             "T", "T", "T", "T", "G", "G", "G", "G", "G", "G", "G", "G", "G", "G"),
    length = c("S", "S", "S", "S", "S", "L", "L", "L", "L", "L", "S", "S", "S", 
               "S", "S", "L", "L", "L", "L", "L", "S", "S", "S", "S", "S", "L", 
               "L", "L", "L", "L", "S", "S", "S", "S", "S", "L", "L", "L", "L", "L"),
    pref = c("F", "I", "R", "B", "O", "F", "I", "R", "B", "O", "F", "I", "R", 
             "B", "O", "F", "I", "R", "B", "O", "F", "I", "R", "B", "O", "F", 
             "I", "R", "B", "O", "F", "I", "R", "B", "O", "F", "I", "R", "B", "O"),
    count = c(23, 4, 2, 2, 8, 7, 0, 1, 3, 5, 5, 11, 1, 0, 3, 13, 8, 6, 1, 0, 5, 
              11, 2, 1, 5, 8, 7, 6, 3, 5, 16, 19, 1, 2, 3, 17, 1, 0, 1, 3)
  )

gator$lake = relevel(gator$lake, ref = "H")
gator$length = relevel(gator$length, ref = "S")
gator$pref = relevel(gator$pref, ref = "O")

```

```{r n, comment=NA}
(mdl = multinom(pref ~ lake + length, weights = count, data = gator))
```

a. $log(\frac{\pi_B}{\pi_O}) = -1.28 + .131(lake_G) - .527(lake_O) - .297(lake_T) + .962(length_L)$  
   $log(\frac{\pi_F}{\pi_O}) = .7465 + .826(lake_G) + .820(lake_O) - .690(lake_T) + .332(length_L)$  
   $log(\frac{\pi_I}{\pi_O}) = -1.00 + 2.48(lake_G) + 3.41(lake_O) + 2.09(lake_T) - 1.12(length_L)$  
   $log(\frac{\pi_R}{\pi_O}) = -1.67 - .416(lake_G) + 2.04(lake_O) + 1.00(lake_T) + .682(length_L)$  


b. If the length is short the probability that the gator will prefer Fish is .258, if the gator is long then the probability that it will prefer fish is .458

```{r o, comment=NA}
predict(mdl, data.frame(lake = c("O", "O"), length = c("S", "L")), type = "probs")
```

\newpage

### 6.4

```{r p, comment=NA, echo=FALSE}
afterlife = data.frame(
  response = rep(c("Yes", "Undecided", "No"), 4),
  race = c(rep("White", 6), rep("Black", 6)),
  gender = rep(c(rep("Female", 3), rep("Male", 3)), 2),
  count = c(371, 49, 74, 250, 45, 71, 64, 9, 15, 25, 5, 13)
)
afterlife$gender = relevel(afterlife$gender, ref = "Female")
afterlife$race = relevel(afterlife$race, ref = "White")
summary(afterlife)
```

a. $\pi_Y = exp(1.64 - .341) / (1 + exp(1.64 - .341) + exp(-.382 - .27)) = .707$

```{r q, comment=NA}
(mdl = multinom(response ~ race + gender, weights = count, data = afterlife))
predict(mdl, data.frame(gender = "Female", race = "Black"), type = "probs")
```

b. $log(\frac{\pi_U}{\pi_N}) = -.382 - .27(Race_B) - .105(Gender_M)$ The conditional odds of Undecided to No is multiplied by exp(-.105) = .9 when the gender is male vs female
$log(\frac{\pi_Y}{\pi_U}) = (-.382 - 1.643) - (.2709 - .341)Race_B - (.105 - .418)Gender_M$ The conditional odds ratio of Yes to Undecided is multiplied by exp(-.105 - .418) = .592 when gender is male.

\newpage

### 6.5

a.
  i.  $x_1$: Those who make more money relative to their peers are less satisfied with their job compared to those who make less money `exp(-.54 * 1:4) / (1 + exp(-.54 * 1:4)) = [1] 0.3681876 0.2535060 0.1652049 0.1034005`
  ii. $x_2$: Employees have higher job satisfaction with less freedom `exp(.6 * 1:4)  / (1 + exp(.6 * 1:4)) = [1] 0.6456563 0.7685248 0.8581489 0.9168273`
  ii. $x_3$: Work environments that are less productive have higher job satisfaction `exp(1.19 * 1:4) / (1 + exp(1.19 * 1:4)) = [1] 0.7667411 0.9152894 0.9726152 0.9915071`

b. x1 = 1, x2 = 4, x3 = 4


### 6.6

a. $log(\frac{\pi_{H1}}{\pi_{H3}}) = -2.55 - .2275x$  
   $log(\frac{\pi_{H2}}{\pi_{H3}}) = -.353 - .0962x$  
   $log(\frac{\pi_{H1}}{\pi_{H2}}) = (-2.3 - .353) - (.2275 - .0962)x$

b. The increase in odds for marital happiness increases exp(-.2275) = .79 for each increased level of income

c. Wald: .9432, Pvalue: .624

d. Based on the value/df ratio of 1.5 it appears that the mode does not fit adequately. The ratio should be close to 1 with a well fit model.

e. $\pi_3 = 1 / (1 + exp(-2.55 - .2275) + exp(-.351 - .0962)) = .587$


### 6.7

a. 2 intercepts are fit because the response has 3 levels. There is only 1 income effect because it is assumed fixed across the 3 levels. If it were not fixed then there is no garuntee that the sum of the 3 probabilities of the response would add to 1.
b. For each unit increase in income, odds of of increasing marital hapiness is multiplied times exp(.65) = 1.91
c. Chi-square: .897, pvalue: .343, because the pvalue is > .05 we conclude that income does not have a significant affect on marital happiness
d. Based on the deviance/df = 1.08, the model fits the data reasonably well.
e. $\pi_2 = \frac{exp(-.237 + (2 * -.117)) - exp(-3.24 + (2 * -.111))}{(1 + exp(-3.24 + (2 * -.111)))*(1 + exp(-.237 + (2 * -.117)))} = .35$
 
\newpage

### 6.8

a. Sequential Effect: $exp(.5807) = 1.78$ The increase in odds for Sequential over Alternating is a multiplier of 1.78 
   Gender Effect: $exp(.541) = 1.71$ The increase in odds for Male over Female is a multiplier of 1.78  

```{r s, comment=NA, echo=FALSE}

cancer = data.frame(
  therapy = c(rep("Sequential", 8), rep("Alternating", 8)),
  gender = rep(c(rep("Male", 4), rep("Female", 4)), 2),
  response = rep(c("Progressive", "No Change", "Partial Remission", "Complete Remission"), 4),
  count = c(28, 45, 29, 26, 4, 12, 5, 2, 41, 44, 20, 20, 12, 7, 3, 1)
)

cancer$response = 
  factor(cancer$response,
         levels = c("Progressive", "No Change", "Partial Remission", "Complete Remission"),
         ordered = TRUE)

summary(cancer)

```

```{r s2, comment=NA}

library(MASS)
mdl = polr(response ~ therapy + gender, weights = count, data = cancer); summary(mdl)

```


\newpage

b. Interaction: $exp(-.59) = .55$ The treatment effect for a Male vs a female is a multiplier of .55 so the treatment is more effective for females than it is for males.

```{r s3, comment=NA}

mdl = polr(response ~ therapy*gender, weights = count, data = cancer); summary(mdl)

```

c. Based on the lower AIC the interaction model is better than the model with no interaction


### 6.9

a. Political ideology is an ordered factor so you expect the coefficients to be ordered as well. There are only 4 because the 5th logit is redundant and can dervied from the other 4.
b. The group by that is the most liberal is the relgious group: none because all of the other coefficients for religion are negative. The group that is the most conservative is the religion group: Protestant because the have the smallest coefficient.
c. Protestant: $exp(-1.03 - 1.27) / (1 + exp(-1.03 - 1.27)) = .091$  
   None: $1 / (1 + exp(-1.03 - .44)) = .81$
d. Protestant/None: $exp(-1.27) = .28$ The effect of being more conservative between Protestant/None is a multiplier of .28
   Protestant/Catholic: $exp(-1.27 - 1.22) = .082$ The effecto of being more conservative between Protestant/Catholic is .082
   
   
Additional

$log(\frac{\pi_1}{\pi_3}) = \alpha_1 + \beta_1 x \rightarrow \frac{\pi_1}{\pi_3} = exp(\alpha_1 + \beta_1x) \rightarrow \pi_1 = \frac{exp(\alpha_1 + \beta_1x)}{(1 + exp(\alpha_1 + \beta_1x)) + (1 + exp(\alpha_2 + \beta_2x))}$  
$log(\frac{\pi_2}{\pi_3}) = \alpha_2 + \beta_2 x \rightarrow \frac{\pi_2}{\pi_3} = exp(\alpha_2 + \beta_2x) \rightarrow \pi_2 = \frac{exp(\alpha_2 + \beta_2x)}{(1 + exp(\alpha_1 + \beta_1x)) + (1 + exp(\alpha_2 + \beta_2x))}$  
$\pi_3 = 1 - \pi_1 - \pi_2$