---
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: Calibri Light
sansfont: Calibri Light
fontsize: 10pt
geometry: margin=1in
---

**Homework 11**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 659-700**  


### 8.2

a) We reject the null hypothesis and conclude that the marginal probabilities of answering Yes are different. The 95% confidence interval for the difference in the two proportions is:

```{r a, comment=NA, warning=FALSE, message=FALSE}
N = 833 + 125 + 2 + 160

p.1p = (833 + 125) / N; p.p1 = (833 + 2) / N

p.1p - p.p1 + c(-1, 1) * 1.96 * sqrt( ((125 + 2) - ((125 - 2)^2 /N)) / N^2)

```

b) 90% confidence interval for the difference in proportions

```{r a2, comment=NA, warning=FALSE, message=FALSE}
p.1p - p.p1 + c(-1, 1) * 1.645 * sqrt( ((125 + 2) - ((125 - 2)^2 /N)) / N^2)

```

### 8.3

\begin{align*}
  logit[P(Y_2 = 1)] = & log(\frac{(835/1120)}{1 - (835/1120)}) = 1.075 \\
  \beta = & log(\frac{(958/162)}{(835/285)}) = .702 \\
  \text{Marginal Model} = & 1.075 + .702x \\
\end{align*}

a) The odds that someone believes in both heaven and hell is exp(.702) = 2.02 times greater than someone who does not believe in heaven and hell

b) The odds that someone believes in hell conditional that they believe in heaven is exp(1.075 + .702) = 5.91 times the odds that someone believes in heaven, conditional that they believe in hell

### 8.5

a) If the probabilities of "No" for both questions is the same then probability of a response being a No in one question, but not in the other is .5.

b) The mid pvalue is 1/2 the probability of one more "No" response plus the probability of "No" for the same question.

c) The binomial distribution is discrete and so it cannot calculate exact probabilities. The Mid pvalue is used as a correction to try to get closer to the exact pvalue.

\newpage

### 8.6

a) We reject the null hypthothesis and conclude that there is a significant difference in proportions

```{r b,  comment=NA, warning=FALSE, message=FALSE}
mcnemar.test(x = matrix(c(114, 11, 181, 48), nrow = 2))
```

b) The 90% confidence interval for difference in proportions of a female opposing info and opposing health opinions is between .43 and .53

```{r b2,  comment=NA, warning=FALSE, message=FALSE}
N = 114 + 181 + 11 + 48

p.1p = (114 + 181) / N; p.p1 = (114 + 11) / N

p.1p - p.p1 + c(-1, 1) * 1.645 * sqrt( ((181 + 11) - ((181 - 11)^2 /N)) / N^2)

```

c) The odds ratios for Supporting Info Opinion is exp(2.214) = 9.152 times higher than Supporting Health Opinion. The conditional odds ratio of Supporting Info Opinion is exp(-.605 + 2.214) = 4.99 times higher than Supporting Health Opinion.

\begin{align*}
  logit[P(Y_2 = 1)] = & log(\frac{125/354}{1 - (125/354)}) = -.605 \\
  \beta = & log(\frac{295/59}{125/229}) = 2.214 \\
  \text{Marginal Model} = & -.605 + 2.214x \\
\end{align*}

d) You could create a new 2x2 table by taking the difference in the counts between Males and Females for each question. 

### 8.7

`(.314 - .292) + c(-1, 1) * 1.96 * sqrt( ((359 + 334) - ((359 - 334)^2 / 1144)) / 1144^2 ) = (-.023, .067)`

The interval is wider than the interval in the book.  The estimate could gain higher precision by adding more observations.

### 8.11

The odds of someone answering yes to cuts in paying higher taxes is .95 to 1.59 times that of those who answer yes to cut in living standard. Because the 95% confidence interval crosses 1, the odds ratio of the two questions are not significantly different.

$log(132/107) \pm * 1.96 * \sqrt(1/132 + 1/107) \rightarrow exp(-.0449, .4649) = [.956, 1.59]$

\newpage

### 8.13

a) $G^2 = 150.6 > 79 = qchisq(.95, 6)$ Reject the null hypthosesis that there is symmetry in the responses
b) $G^2 = 2.3 < 7.8 = qchisq(.95, 3)$ Do not reject the null hypothesis there is quassi symmetry
c) $G^2 = 150.6 - 2.3 = 148.3 > 7.8 = qchisq(.95, 3)$ Reject the null hypothesis that there is marginal homogeneity
d) 


### 8.14

a)

```{r c, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}
r.16=rep(c("ne","mw","s","w"),c(4,4,4,4))
r.2004=rep(c("ne","mw","s","w"),4)
count=c(425,17,80,36,10,555,74,47,7,34,771,33,5,14,29,452)
r.16=factor(r.16)
r.2004=factor(r.2004)

### Create indicator variables for each diagonal cell
r.16_ne=(r.16=="ne")
r.16_mw=(r.16=="mw")
r.16_s=(r.16=="s")
r.16_w = (r.16 == "w")
symm1=1*(r.16=="ne")*(r.2004=="ne")
symm2=2*(r.16=="mw")*(r.2004=="mw")
symm3=3*(r.16=="s")*(r.2004=="s")
symm4=4*(r.16=="mw")*(r.2004=="ne")+4*(r.16=="ne")*(r.2004=="mw")
symm5=5*(r.16=="mw")*(r.2004=="s")+5*(r.16=="s")*(r.2004=="mw")
symm6=6*(r.16=="ne")*(r.2004=="s")+6*(r.16=="s")*(r.2004=="ne")
symm7=7*(r.16=="mw")*(r.2004=="w")+6*(r.16=="w")*(r.2004=="mw")
symm8=8*(r.16=="ne")*(r.2004=="w")+7*(r.16=="w")*(r.2004=="ne")

### Symmetry Model

model=glm(count~symm1+symm2+symm3+symm4+symm5+symm6+symm7,family=poisson(link=log))
summary(model)

### Quasi-Symmetry Model

model=glm(count~r.16+r.2004
          +symm1+symm2+symm3+symm4+symm5+symm6+symm7,family=poisson(link=log))
summary(model)
```

b) $G^2 = 1114 - 125 = 989 > 12.59 = qchisq(.95, 6)$ Reject Null and conclude that there marginal homoegeniety

\newpage

### 8.16

There is evidence of symmetry $G^2 = 364 > 11.07 = qchisq(.95, 5)$. There is also strong evidence of quasi symmetry with $G^2 = 807 > 14 = qchisq(.95, 7)$. Marginal homogeniety is also present with $G^2 = 445 - 1.22 = 443.98 > 6 = qchisq(.95, 2)$.

```{r d, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}
Chem=rep(c("1","2","3"),c(3,3,3))
Recycle=rep(c("1","2","3"),3)
count=c(66,227,150,39,359,216,3,48,108)
Chem=factor(Chem)
Recycle=factor(Recycle)

### Create indicator variables for each diagonal cell
Chem_1=(Chem=="1")
Chem_2=(Chem=="2")
Chem_3=(Chem=="3")
Recy_1=(Recycle=="1")
Recy_2=(Recycle=="2")
Recy_3=(Recycle=="3")
symm1=1*(Chem=="1")*(Recycle=="1")
symm2=2*(Chem=="2")*(Recycle=="2")
symm3=3*(Chem=="3")*(Recycle=="3")
symm4=4*(Chem=="2")*(Recycle=="1")+4*(Chem=="1")*(Recycle=="2")
symm5=5*(Chem=="2")*(Recycle=="3")+5*(Chem=="3")*(Recycle=="2")

### Symmetry Model

model=glm(count~symm1+symm2+symm3+symm4+symm5,family=poisson(link=log))
summary(model)

### Quasi-Symmetry Model

model=glm(count~Chem + Recycle +symm1+symm2+symm3+symm4+symm5,
          family=poisson(link=log))
summary(model)

### Ordinal Quasi-Symmetry Model

model=glm(count~Chem*Recycle+symm1+symm2+symm3+symm4+symm5,
          family=poisson(link=log))
summary(model)

```

\newpage

### 8.17

Deviance for the symmetry model is 40.5 on 6 degrees of freedom. Deviance from the quasi-symmetry model is 27.4 on 5 degrees of freedom $G^2 = 40.5 - 27.4 = 13.1 > 3.84 = qchisq(.95, 1)$ Reject the null and conclude that the symmetry model is not an adequate fit. 

Output from the symmetry model

| Parameter		| DF | 	Estimate	|   SE    |	Wald 95% |   CL   |	Chi-Square |	Pr > ChiSq |
|-------------|----|------------|---------|----------|--------|------------|-------------|
| Intercept		| 1	 | 3.2581	    | 0.1961  | 2.8737	 | 3.6425	| 276	       | <.0001      |
| symm1	 	    | 1  |	1.2958	  | 0.2213	| 0.862	   | 1.7296	| 34.27	     | <.0001      |
| symm2	 	    | 1  |	0.976	    | 0.2138	| 0.557	   | 1.395	| 20.84	     | <.0001      |
| symm3	 	    | 1  |	0.1919	  | 0.2331	| -0.265	 | 0.6488	| 0.68	     | 0.4104      |
| symm4	 	    | 1  |	-1.3863	  | 0.3397	| -2.0521	 | -0.7205| 16.66	     | <.0001      |
| symm5	 	    | 1  |	1.6017	  | 0.215	  | 1.1804	 | 2.0231	| 55.51	     | <.0001      |
| symm6	 	    | 1  |	1.4287	  | 0.2075	| 1.0219	 | 1.8354	| 47.39	     | <.0001      |
| symm7	 	    | 1  |	-1.118	  | 0.3119	| -1.7294	 | -0.5067| 12.85	     | 0.0003      |
| symm8	 	    | 1  |	2.1929	  | 0.2068	| 1.7877	 | 2.5982	| 112.48     | 	<.0001     |
| symm9	 	    | 1  |	0.7122	  | 0.2189	| 0.2833	 | 1.1411	| 10.59	     | 0.0011      |
| symm10 	    | 0  |	0	        |0	      | 0	       | 0	    |  .	       | .           |


### 8.19

Deviance for the independence model 4188 and deviance for the quasi-independence model is 3.7. There is a large improvement in the model fit because the quaisi-independnence model uses an identity matrix to weight the observations in the trace.

```{r e2, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}
options(width = 100)

library(catspec)
r.16=rep(c("ne","mw","s","w"),c(4,4,4,4))
r.2004=rep(c("ne","mw","s","w"),4)
count=c(425,17,80,36,10,555,74,47,7,34,771,33,5,14,29,452)
r.16=factor(r.16)
r.2004=factor(r.2004)

dta = data.frame(
  r.16 = r.16,
  r.2004 = r.2004,
  count = count
)

### Independence Model
model=glm(count~r.16+r.2004, family=poisson(link=log))
model

### Quasi-Independence Model
model=glm(count~r.16+r.2004+mob.symint(r.16, r.2004)+
            mob.qi(r.16, r.2004),
          family=poisson(link=log))
model

```

### 8.20

The quasi symmetry model fits much better than the independence model given that the residual deviance is much lower. The kappa coefficient indicates that the difference between the agreements is .21 that of the maximum possible difference under the independence model. We conclude that there is significant differences in the agreements of the two neurologists.

```{r f,  echo=FALSE, message=FALSE, comment=NA, warning=FALSE}
dta = data.frame(
  A = factor(rep(1:4, rep(4,4))),
  B = factor(rep(1:4, 4)),
  C = c(38,33,10,3,5,11,14,7,0,3,5,3,1,0,6,10)
)

model = glm(C ~ A + B, family = poisson, data = dta)
summary(model)
model = glm(C ~ A + B + mob.symint(A, B), family = poisson, data = dta)
summary(model)

library(psych)
cohen.kappa(xtabs(C ~ A + B, data = dta))
```

\newpage

### 8.22

```{r g, message=FALSE, comment=NA, warning=FALSE}
library(BradleyTerry2)

dta = data.frame(
  win.1 = c(20, 19, 31),
  win.2 = c(29, 28, 19),
  player.1 = c("Pepsi", "Classic Coke", "Coke"),
  player.2 = c("Coke", "Pepsi", "Classic Coke")
)

mdl = BTm(cbind(dta$win.1, dta$win.2), player.1, player.2, id = "drink", data = dta)
summary(mdl)
fitted(mdl)
```

a) Rankings: Coke, Pepsi, Classic Coke
b) 

| Comparison       | Predicted | Sample |
|------------------|-----------|--------|
| Coke vs Pepsi    |    .57    |  .592  |

\newpage

### 8.23

```{r h, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

## Convert frequencies to success/failure data
citations.sf = countsToBinomial(citations)
names(citations.sf)[1:2] = c("journal1", "journal2")

##  First fit the "standard" Bradley-Terry model
citeModel = BTm(cbind(win1, win2), journal1, journal2, data = citations.sf)

citations

summary(citeModel)

fitted(citeModel)
```

a) Prestige Ranking: JRSS-B, Biometrika, JASA, Comm Statist
b) The Probability that the Comm Stat journal cites the JRSS-B journal is .038

### 8.27

a) $log(\mu_{ij} / \mu_{ji}) = log(\mu_{ij}) - log(\mu_{ji}) \rightarrow exp[log(\mu_{ij}) - log(\mu_{ji})] = \mu_{i} - \mu_{j} = \beta_i - \beta_j$

b) $\lambda_i^X = \lambda_j^y \rightarrow \lambda_i^X - \lambda_j^Y = 0 \rightarrow log(\mu_{ij}) = \lambda + \lambda_i + \lambda_j + \lambda_{ij}$

c) $I(i \ne j) = 0 = \lambda_{ij} \rightarrow log(\mu_{ij}) = \lambda + \lambda_i + \lambda_j + \lambda_{ij}$
















