---
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: DejaVu Sans Mono
sansfont: DejaVu Sans Mono
fontsize: 11pt
geometry: margin=1in
---

Homework 01  
Joseph Blubaugh  
jblubau1@tamu.edu  

```{r a, echo=FALSE, results='hide'}
library(knitr)
library(scales)
library(ggplot2)
```


### 2.1

Rows: Fathers Profession, Cols: Sons Profession

```{r b, echo=FALSE}
jobs = data.frame(row.names = c('farm', 'operative', 'craftsman', 'sales', 'professional'),
                  farm = c(.018, .002, .001, .001, .001),
                  operative = c(.035, .112, .066, .018, .029),
                  craftsman = c(.031, .064, .094, .019, .032),
                  sales = c(.008, .032, .032, .01, .043),
                  professional = c(.018, .069, .084, .051, .13)
                  )
kable(jobs)
```

#### a)  Fathers marginal density

```{r c, comment=NA}
rowSums(jobs)
```

#### b)  Sons marginal density

```{r d, comment=NA}
colSums(jobs)
```

#### c) P(Son Profession | Father = Farmer)

```{r e, comment=NA}
jobs[1, ] / rowSums(jobs)[1]
```

#### d) P(Father Profession | Son = Farmer)

```{r f, comment=NA}
data.frame(farm = jobs[, 1] / colSums(jobs)[1], 
           row.names = c('farm', 'operative', 'craftsman', 'sales', 'professional'))
```

\newpage

### 2.2

a)  $E[a_1 Y_1 + a_2 Y_2] =  a_1 E[Y_1] + a_2 E[Y_2]$  
 $Var[a_1 Y_1 + a_2 Y_2] =  a_1^2 Var[Y_1] + a_2^2 Var[Y_2]$  
 
b)  $E[a_1 Y_1 - a_2 Y_2] =  a_1 E[Y_1] - a_2 E[Y_2]$  
 $Var[a_1 Y_1 - a_2 Y_2] =  a_1^2 Var[Y_1] + a_2^2 Var[Y_2]$  

### 2.3

a)  $P(X|Y) = P(X)$; $P(X|Y, Z) = P(X,Z) \approx f(X,Z)$

b)  $P(Y|X) = P(Y)$; $P(Y|X, Z) = P(Y,Z) \approx f(Y,Z)$
 
### 2.5

a)

Rows: X, Cols: Y

```{r g, comment=NA, echo=FALSE}
dta = data.frame(
  row.names = c("Head", "Tail", "Marginal.Y"),
  "Green" = c(.2, .3, .5),
  "Red" = c(.3, .2, .5),
  "Marginal.X" = c(.5, .5, 1)
)

kable(dta, format = "markdown")

```

b)  $E[Y] = .5$  
    $P[Y = G | X = H] = .2$  
    $P[Y = G | X = T] = .3$  
    $P[Y = G] = .5$
    
c)  $Var[Y|X = 0] = 0^2 (.2) + 1^2 (.3) = .3$  
    $Var[Y|X = 1] = 0^2 (.3) + 1^2 (.2) = .2$  

The difference in variance is because we are essentially comparing the following conditional probabilities: $P(Y = 1 | X = 0) = .3$ and $P(Y = 1 | X = 1) = .2$. The smaller probability will have smaller variation than the larger probability because we are more certain that the smaller probability will occur less often than the larger probability.

d)  $P(X = 0 | Y = 1) = .2/.5 = .4$

\newpage

### 2.6

Under this scenario the following are conditionally dependent. $P(A \bot B | C); P(A^c \bot B | C); P(A \bot B^c | C)$ and $A^c, B^c$ are the same area. $P(A \bot B | C^c)$ are not conditionally independent because A and B intersect outside of C.  

```{r h, echo=FALSE}
ggplot() +
  geom_tile(aes(x = 2, y = 4), width = 5, height = 5, fill = "yellow", alpha = .3, color = "black", lty = 2) +
  geom_tile(aes(x = 5, y = 4.5), width = 1, height = 4, fill = "yellow", alpha = .3, color = "black", lty = 2) +
  geom_tile(aes(x = 3, y = 1), width = 7, height = 3, fill = "blue", alpha = .3, color = "black") +
  geom_tile(aes(x = 5.5, y = 3), width = 2, height = 7, fill = "red", alpha = .3, color = "black") +
  annotate("text", x = 0, y = 6, label = "C") +
  annotate("text", x = 0, y = 0, label = "A") +
  annotate("text", x = 6, y = 6, label = "B") +
  annotate("text", x = 6, y = 0, label = "A n B") +
  annotate("text", x = 0, y = 2, label = "A n C") +
  annotate("text", x = 5, y = 6, label = "B n C") +
  annotate("text", x = 5, y = 2, label = "C^c") +
  annotate("text", x = 4, y = 1, label = "C^c") +
  annotate("text", x = 6, y = 3, label = "C^c") +
  theme_void()


```













