---
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: Calibri Light
sansfont: Calibri Light
fontsize: 11pt
geometry: margin=1in
---

**Homework 08**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 641-720**  

\newpage

1) Type I Error: $P(z > \frac{\sqrt{10} 500}{1000}) = P(z > \frac{\sqrt{10}}{2}) = .056$  
Type II Error: $P(z \le \frac{-200 \sqrt(10)}{1000}) = P(z \le \frac{-\sqrt{10}}{5}) = .263$

2) 
    a)  Decision Rule $Z = \frac{\sqrt{n} (\bar{Y} - \mu_0)}{\sigma} > Z_{1 - \alpha} = \frac{\sqrt{10} (\bar{Y} - 10500)}{1000} > 2.326$  
    We reject the null hypothesis when $\bar{Y} > 11235.65$> 
    b)  

```{r q2b, comment=NA}
alpha = qnorm(.99)
mu = 10500 + alpha * 1000/sqrt(10)
x = c(10600, 10800, 11000, 11500)
pnorm((sqrt(10) * (x - mu))/1000)
```

3) 
    a) p-value: `r pnorm((sqrt(15) * (8.7 - 10))/2)`, we conclude that there is sufficient evidence to reject the null hypothesis because the pvalue is less than .01
\begin{align*}
    H_o: \mu \ge 10, H_1: \mu < 10 \\
    \operatorname{reject} H_o: \operatorname{when} \bar{y} < 10 - Z_{.01} \frac{2}{\sqrt{15}} = 8.79 \\
    \bar{Y} = 8.7 \\
    \operatorname{}
\end{align*}
    b)$H_o: \mu > 8.5, H_1: \mu \le 8.5$

```{r q3b, comment=NA}
alpha = -qt(.99, 14)
delta = (sqrt(15)*(.2))/2
pt(alpha, 14, delta)
```

    c)


4)

```{r q4, comment=NA}
dt = c(125,123,117,123,115,112,128,118,124,
       111,116,109,125,120,113,123,112,118,
       121,118,122,115,105,118,131)

# a)
(t = 5 * (10 - sd(dt)) / 10); pnorm(t)

# b)
sigma = c(5, 6, 7, 8, 9, 10)
pnorm(qnorm(.9) + (5 * (sigma - sd(dt)) / sigma))

# c) This is consistent with the results from a), we will reject the null
(T = qt(.9, 24))

```

5)

```{r q5, comment=NA}
# a) There sufficient evidence to support the alternative hypothesis that the median is less than 120
alpha = .05
n = 21
Splus = qbinom(alpha, n, .5)
pbinom(Splus, n, .5)

# b) There is not significant evidence that the true median is less than 120
wilcox.test(x = dt, y = rep(120, 25), paired = TRUE,
            alternative = "less", conf.level = .95)

# c) lower bound with alpha = .05
qbinom(.05, 21, .5)

```

6)

```{r q6, comment=NA}
# a)
binom.test(46, 50, p = .8, alternative = "greater", conf.level = .95)

# b) There is substantial evidence that the improved method has increased the accuracy over the current method

# c) The accuracy is 90%
Y = qbinom(.95, 50, .8)
1 - pbinom(Y-1, 50, .92)

# d) n = 49 would result in a power of 80%

```


MC1) C

MC2) C

```{r mc2, comment=NA}
n = c(140, 100, 98, 35)
pnorm(sqrt(n)*(1.5)/9)
```

MC3) A

```{r mc3, comment=NA}
# sigma can be any number
sigma = 3; delta = 17 + .5*sigma
((sigma * (qnorm(.05) + qnorm(.1)))/(17 - delta))^2
```

MC4) C

MC5) A

```{r mc5, comment=NA}
1 - pt(qt(.95, 9), 9, 0)
```

MC6) C

MC7) C

MC8) B

MC9) B

MC10) C
