---
title: ''
output:
  pdf_document:
    highlight: tango
---
##### Joseph Blubaugh
##### Hw06, STAT 630


##### 1) 3.5.4
  a) -4(1/11 / 2/11) + 6(1/11 / 2/11) = -2 + 3 = 1
  b) -4(2/11 / 3/11) + 6(1/11 / 3/11) = -8/3 + 2 = -2/3
  c) -4(4/11 / 5/11) + 6(1/11 / 5/11) = -16/5 + 6/5 = -2
  d) -4(0) + 6(1/11 / 1/11) = 6
  e) 1 + -2/3 - 2 + 6 = 4.333

##### 2) 3.5.11ace
  a) 
    \begin{align*}
    E(x) &= \frac{6}{19} \int_0^2 \int_0^1 x^2 + y^3 \, dydx \\
         &= \frac{6}{19} \int_0^2 x^2 y + \frac{y^4}{4} |_0^1 \, dx \\
         &= \frac{6}{19} \int_0^2 x (x^2 + \frac{1}{4}) \, dx \\
         &= \frac{6}{19} (\frac{x^4}{4} + \frac{x^2}{8} |_0^2) \\
         &= 4.5 \\
    \end{align*}
  c)
    \begin{align*}
      E(x|y) &= \frac{57}{96} \int_0^2 x^2 y^{-3} \, dx \\
             &= \frac{57}{96} (\frac{x^3}{3} y^-3 |_0^2) \\
             &= \frac{1.58}{y^3}
    \end{align*}
  e)

##### 3) 3.5.16
  \begin{align*}
    E(x) &= \int x f(x) \, dx \\
    \\
    F_{x|y}(x|y) &= \frac{\lambda^a x^{a - 1} / \gamma (a)}{e^{-x/\lambda}/\lambda} \\
    &= \frac{\lambda \gamma (a)}{e^{-x / \lambda} \lambda^a x^{a - 1}} \\
  \end{align*}

##### 4) 3.6.3
  a) P(x > 9) = E(x) = 1, 1/9 = .111
  b) P(x > 2) = 1/2
  c) P(|x - 1| > 1) = 2 / 1^2 = 2, but this extends outside of the probability bounds and doesn't tell us anything
  d) b is equal to the mean of the distribution and c is equal to the variance
  e) P(x > 9) = (1 - .5)^9 * .5 = .0009,
     P(x > 2) = (1 - .5)^2 * .5 = .25,
     P(|x - 1| > 1) = .5

##### 5) 3.6.11
  a) $E(z) = \int_0^2 z \frac{z^3}{4} \, dz = \frac{z^5}{20} |_0^2 = 1.6$
  b)
    \begin{align*}
      E(z^2) &= \int_0^2 z^2 f(x) \, dx \\
      &= int_0^2 \frac{z^5}{4} \, dz \\ 
      &= \frac{z^6}{24} |_0^2 \\
      &= 2.66 \\
      \\
      Var(z) &= 2.66 - 1.6 \\
      &= 1.06 \\
      &= \frac{1}{1.06} \\
      &= .94 \\
    \end{align*}

##### 6) additional a
  $\lambda \theta$ is a poisson distribution because the following holds true when using the mgf to find the mean and variance
  \begin{align*}
    with mgf \\
    f(t) &= e^{\lambda \theta (e^t - 1)} \\
    f'(t) &= \lambda \theta e^{\lambda \theta (e^t -1) + t} \\
    f'(0) &= \lambda \theta \\
    \\
    f''(t) &= \lambda \theta (e \lambda \theta + 1)^{e \lambda \theta (e x−1) + x}
    f''(0) &= \lambda \theta \\
  \end{align*}

##### 7) additional b
  \begin{align*}
    T ~ exponential(\lambda), conditional on T \\
    U ~ uniform[0, T] \\
    \\
    E[E(U|T)] &= E(U) \\
    mean &= \frac{T}{2} \\
    variance &= \frac{T^2}{12}
  \end{align*}

##### 8) 4.1.11
```{r, q8}
# a
max.norm = as.numeric()
for (i in 1:10){max.norm = c(max.norm, max(rnorm(10^6, mean = 0, sd = 1)))}
max.norm
mean(max.norm)
sd(max.norm)

# b) The histogram changes dramatically with only 10 total samples. 
# It looks as though it will be normally distributed given more samples
hist(max.norm)

# c
max.norm.20 = as.numeric()
for (i in 1:20){max.norm.20 = c(max.norm.20, max(rnorm(10^6, mean = 0, sd = 1)))}
max.norm.20
mean(max.norm.20)
sd(max.norm.20)

# d) The histogram is starting to take the shape of a normal distribution
hist(max.norm.20)
```

##### 9) 4.2.2
  \begin{align*}
    ln(X_n) &= n \\
    ln(1) &= 0 \\
    lim_{n \to \infty} X_n &= 0 \\
  \end{align*}

##### 10) 4.2.10
  $1^2 + 2^2 + 3^2 + 4^2 + 5^2 + 6^2 = \frac{91}{6} = 15.1$  
  For large n, the average is very close to 15.1

##### 11) 4.2.12
```{r, q10}
results = as.numeric()
for (i in 1:10^4) results = c(results, mean(rexp(20,5)))
hist(results)

n20 = length(which(results >= .19 & results <= .21)) / length(results)
# The proportion of results that are between .19 and .21 are:
n20

results = as.numeric()
for (i in 1:10^4) results = c(results, mean(rexp(20,5)))
hist(results)

n50 = length(which(results >= .19 & results <= .21)) / length(results)
# Repeated with n = 50, the proportion of results that are between .19 and .21 are:
n50

```

##### 12) 4.4.4
  \begin{align*}
    W_n(x) &= \frac{1 + \frac{x}{n}}{1 + \frac{1}{2n}} \\
    &= \int_0^{\infty} &= W_n(x) \, dx \\
    &= lim_{n \to \infty} \frac{\frac{x^2}{2n} + x}{\frac{n}{2} + 1} \\
    &= x = 1 \\
  \end{align*}

##### 13) 4.4.6
```{r q13}
func = function(x) 1/x * sum(x/(-10))
results.mean = as.numeric()
results.sum = as.numeric()
for (i in 1:10000) {
  results.mean[i] = mean(func(sample(seq(-20,10,.01) , 900, replace = T)))
  results.sum[i] = sum(func(sample(seq(-20,10,.01) , 900, replace = T)))
  }
hist(results.mean, breaks = 40)
hist(results.sum, breaks = 40)

length(which(results.sum > -4470))/length(results.sum)
```

##### 14) 4.4.12
  a) for n = 16, $\frac{2.5 - 2}{2 \sqrt{16}} = \frac{.5}{8} = .062 z-score, 1 - .4761 = .523$
  b) for n = 36, $\frac{2.5 - 2}{2 - sqrt{36}} = \frac{.5}{12} = .04 z-score, 1 - .484 = .516$
  c) for n = 100, $\frac{2.5 - 2}{2 - sqrt{100}} = \frac{.5}{20} = .025 z-score = 1 - .492 = .508$

##### 15) 4.4.16
```{r q15}
func = function(x) 1/x * sum(x/(-10))
results.mean = as.numeric()
results.sum = as.numeric()
for (i in 1:10000) {
  results.mean[i] = mean(func(sample(seq(-20,10,.01) , 30, replace = T)))
  results.sum[i] = sum(func(sample(seq(-20,10,.01) , 30, replace = T)))
  }
hist(results.mean, breaks = 40)
hist(results.sum, breaks = 40)

length(which(results.sum < -5))/length(results.sum)

```

##### 16) 4.5.14
```{r q16}
results = as.numeric()
n = 10000
t = function(U) cos(U^3) * sin(U^4)

for (i in 1:1000) {
  U = rnorm(n)
  round = mean(t(U))
  results = c(results, round)
  }

hist(results, breaks = 40)

mean(results)
mean(results) - sd(results) * 1.96   ## lower 95% confidence
mean(results) + sd(results) * 1.96   ## upper 95% confidence
```






















