---
output:
  pdf_document:
    latex_engine: xelatex
    highlight: pygments
mainfont: Calibri Light
sansfont: Calibri Light
fontsize: 11pt
geometry: margin=1in
---

**Homework 04**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 641-720**  

\newpage

```{r data, echo=FALSE, comment=NA, warning=FALSE}
small = c(0.42,0.86,0.88,1.11,1.34,1.38,1.42,1.47,1.63, 1.73,2.17,2.42,2.48,2.74,2.74,2.79,2.90,3.12,
    3.18,3.27,3.30,3.61,3.63,4.13,4.40,5.00,5.20, 5.59,7.04, 7.15,7.25,7.75,8.00,8.84,9.30,9.68,
    10.32,10.41,10.48,11.29,12.30,12.53,12.69,14.14,14.15, 14.27,14.56,15.84,18.55,19.73,20.00)

large = c(0.94,1.26,1.44,1.49,1.63,1.80,2.00,2.00,2.56, 2.58,3.24,3.39,3.53,3.77,4.36,4.41,4.60,4.67,
    5.39,6.25,7.02,7.89,7.97,8.00,8.28,8.83,8.91, 8.96,9.92,11.36,12.15,14.40,16.00,18.61,18.75,19.05,
    21.00,21.41,23.27,24.71,25.00,28.75,30.23,35.45,36.35)

```

I.  
1) mean = `r round(mean(large),3)`  
   trimmed mean = `r round(mean(large[5:41]),3)` with $K(\alpha) = 45 - 45(.1) - (45(.1) + 1) - 1 = .1$  
   This suggests that there are some extreme values in the top 10 percent that are pulling the mean up

2) $Weibull(\gamma, \alpha) = \frac{\gamma}{\alpha} (\frac{x}{\alpha})^{\gamma - 1} e^{-(x/\alpha)^{\gamma}}$  
  parameter estimates: $\gamma = 1.132, \alpha = 11.498$

```{r q2, echo=FALSE, warning=FALSE, comment=NA}
large = c(0.94,1.26,1.44,1.49,1.63,1.80,2.00,2.00,2.56,
    2.58,3.24,3.39,3.53,3.77,4.36,4.41,4.60,4.67,
    5.39,6.25,7.02,7.89,7.97,8.00,8.28,8.83,8.91,
    8.96,9.92,11.36,12.15,14.40,16.00,18.61,18.75,19.05,
    21.00,21.41,23.27,24.71,25.00,28.75,30.23,35.45,36.35)

small = c(0.42,0.86,0.88,1.11,1.34,1.38,1.42,1.47,1.63,
    1.73,2.17,2.42,2.48,2.74,2.74,2.79,2.90,3.12,
    3.18,3.27,3.30,3.61,3.63,4.13,4.40,5.00,5.20,
    5.59,7.04,7.15,7.25,7.75,8.00,8.84,9.30,9.68,
    10.32,10.41,10.48,11.29,12.30,12.53,12.69,14.14,14.15,
    14.27,14.56,15.84,18.55,19.73,20.00)

fitdistr(x = large, densfun = "weibull")
```
    
3) Assuming $Weibull(\gamma = 1.1326673, \alpha = 11.4982087)$  
   P(x > 30) = `r round(1 - pweibull(q = 30, shape = 1.1326673, scale = 11.4982087),4)`  

4)  

```{r q4, echo=FALSE, comment=NA}
paste("MLE Mean Estimate:", round(mean(rweibull(100000, shape = 1.1326673, scale = 11.4982087)),3),
      "  MLE Sigma Estimate:", round(sd(rweibull(100000, shape = 1.1326673, scale = 11.4982087)),3))

paste("Distribution Free Mean:", round(mean(large),3), "  Distribution Free Sigma:", round(sd(large),3))
```

5)

```{r q5, echo=FALSE, comment=NA}
paste("MLE Median Estimate:", round(median(rweibull(100000, shape = 1.1326673, scale = 11.4982087)),3),
      "  MLE IQR Estimate:", round(IQR(rweibull(100000, shape = 1.1326673, scale = 11.4982087)),3))

paste("Distribution Free Median:", round(median(large),3), "  Distribution Free IQR:", round(IQR(large),3))
```

6)  

```{r q6, echo=FALSE, comment=NA}
paste("Large Mean:", round(mean(large),3), "  Large SD:", round(sd(large),3))
paste("Small Mean:", round(mean(small),3), "  Small SD:", round(sd(small),3))
```

7)  

```{r q7, echo=FALSE, comment=NA}
paste("Large Median:", round(median(large),3), "  Large MAD:", round(mad(large),3))
paste("Small Mean:", round(median(small),3), "  Small MAD:", round(mad(small),3))
```


8)  Since both distributions are right skewed, median and MAD do the best job at describing the center and spread of the distributions.


9)  Large litters on average have higher brain_wt/body_wt based on the centeroid measurements like median and mean, but large litters also have a much larger variation in brain_wt/body_wt where as the small litters brain_wt/body_wt mass is more concentrated.

II.  

1)

```{r q21, echo=FALSE, comment=NA}
ts = data.frame(dt = c("1989-01-01","1989-02-01","1989-03-01","1989-04-01","1989-05-01",
                       "1989-06-01","1989-07-01","1989-08-01","1989-09-01","1989-10-01",
                       "1989-11-01","1989-12-01","1990-01-01","1990-02-01","1990-03-01",
                       "1990-04-01","1990-05-01","1990-06-01","1990-07-01","1990-08-01",
                       "1990-09-01","1990-10-01","1990-11-01","1990-12-01","1991-01-01",
                       "1991-02-01","1991-03-01","1991-04-01","1991-05-01","1991-06-01",
                       "1991-07-01","1991-08-01","1991-09-01","1991-10-01","1991-11-01",
                       "1991-12-01","1992-01-01","1992-02-01","1992-03-01","1992-04-01",
                       "1992-05-01","1992-06-01","1992-07-01","1992-08-01","1992-09-01",
                       "1992-10-01","1992-11-01","1992-12-01","1993-01-01","1993-02-01",
                       "1993-03-01","1993-04-01","1993-05-01","1993-06-01","1993-07-01",
                       "1993-08-01","1993-09-01","1993-10-01","1993-11-01","1993-12-01"),
                value = c(9.62,9.64,9.80,9.79,9.57,9.10,8.93,8.96,9.01,8.92,8.89,8.96,
                          8.99,9.22,9.37,9.46,9.47,9.26,9.24,9.41,9.56,9.53,9.30,9.05,
                          9.04,8.83,8.93,8.86,8.86,9.01,9.00,8.75,8.61,8.55,8.48,8.31,
                          8.20,8.29,8.35,8.33,8.28,8.22,8.07,7.95,7.92,7.99,8.10,7.98,
                          7.91,7.71,7.58,7.46,7.43,7.33,7.17,6.85,6.66,6.67,6.93,6.90))

ts$dt = as.Date(ts$dt)
plot(ts, type = "l", main = "Daily Yields of Moody AAA bonds")

```

2)  You can conclude that there is a strong positive correlation between adjacent values

```{r 22, echo=FALSE, comment=NA}
acf(ts$value, plot=FALSE)
```

3)  No, it is easy so see that the mean has been declining over the course of time.   

III.  

1)  

```{r 31, echo=FALSE, warning=FALSE, comment=NA}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(survival))

G1 = data.frame(Time = c(180, 632, 2240, 195, 76, 70, 13, 1990, 18, 700, 210, 1296, 23),
                Status = c(1, 1, 1, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1))
                  
G2 = data.frame(Time = c(8, 852, 52, 220, 63, 8, 1976, 1296, 1460, 63, 1328, 365),
                Status = c(0, 1, 1, 1, 1, 1, 0, 0, 1, 1, 1, 1))


survive1 = survfit(Surv(G1$Time, G1$Status) ~ 1, conf.type = "log-log")
survive2 = survfit(Surv(G2$Time, G2$Status) ~ 1, conf.type = "log-log")

print(survive1, print.rmean = TRUE)
summary(survive1)

print(survive2, print.rmean = TRUE)
summary(survive2)

plot(survive1, col = "blue", conf.int = FALSE, log = TRUE, xlab = "Treatment 1 (Blue), Treatment 2 (Red)")
lines(survive2, col = "red", conf.int = FALSE, log = TRUE)

```

2)  Treatment 1: Mean = 657, Median = 202 ; Treatment 2: Mean = 731, Median = 365

3)  Treatment 1 seems to be more effective than treatment 2.

IV.  

1)  Type I censoring, the termination point of the experiment is fixed at 500 psi.  

2)  Left censored, we have complete data because all of the puppies already knew how to swim.

3)  Uncensored, all puppies were able to learn how to swim so all data could be rec

4)  Random censoring, all puppies in this group failed to learn to swim during the time of the study

V.  In this case SIQR = IQR/2
    
\begin{align*}
    r &= \tilde{\mu} - Q(.25)                                    \\
      &= Q(.75) - \tilde{\mu}                                    \\
                                                                 \\
 SIQR &= \frac{r + \tilde{\mu} - \tilde{\mu} - r}{2}             \\
      &= r                                                       \\
                                                                 \\
 P \Big [ Q(.25) \le y \le Q(.75) \Big ] &= .5                   \\
 P \Big [ \tilde{\mu} - r \le y \le \tilde{\mu} + r \Big ] &= .5 \\
 P \Big [ -r \le (y - \tilde{\mu} \le r) \Big ] &= .5            \\
 P \Big [ |(y - \tilde{\mu} | \le r) \Big ] &= .5                \\
 r = SIQR &= |y - \tilde{\mu}|                                   \\
\end{align*}
