---
output:
  pdf_document:
    fig_height: 5
    fig_width: 8
    highlight: pygments
    latex_engine: xelatex
mainfont: Calibri
sansfont: Calibri
fontsize: 12pt
geometry: margin=1in
---

**Homework 06**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 636-720**  

\newpage


1)

```{r q1, warning=FALSE, comment=NA, message=FALSE}
## a)
X1 = read.delim("T8-4.DAT", header=FALSE)
colnames(X1) = c("x1", "x2", "x3", "x4", "x5")

head(X1)

## Sample Variance
S1 = var(X1)

## Pricipal Components
(pca.1 = prcomp(X1))

## Multiply the data with the eigen vectors
PC.1 = as.matrix(X1) %*% eigen(S1)$vectors
head(PC.1, 5)

## b)
## Proportion of variance explained by PCA
diag(var(PC.1)) / sum(diag(S1)); summary(pca.1)

```

In the first component all of the variables have the same sign and so it can be interpreted as a weighted average linear combination.  The ammount of total variance explained by the first component is 53% with the first 3 explaining 90% of the variance in the data. Depending on the application PC4 and PC5 may not be useful since they do not explain the variance very well.

\newpage

2)
a)
```{r q2, warning=FALSE, comment=NA, message=FALSE}

X2 = read.delim("T8-5.DAT", header=FALSE)
colnames(X2) = c("x1", "x2", "x3", "x4", "x5")

## Sample Variance
S2 = var(X2)

## PCA on Variance Matrix
(pca.2s = prcomp(X2))
summary(pca.2s)

PCA.2s = as.matrix(X2) %*% pca.2s$rotation
round(cor(X2, PCA.2s), 3)

##confirm correlation
(eigen(S2)$vectors[1,1] * sqrt(eigen(S2)$values[1])) / sqrt(S2[1,1])

## PCA on Correlation Matrix
(pca.2r = prcomp(X2, center = TRUE, scale. = TRUE))
summary(pca.2r)

PCA.2r = as.matrix(X2) %*% pca.2r$rotation
round(cor(X2, PCA.2r), 3)

```

b)

##### PCA.2s
93% of the variation can be explained by the first 2 pricipal components. The first components looks primarily to be made up of the weighted differences between x3 anx x4. The second component looks to be a weighted average of x2, x3, and x4.

##### PCA.2r
67% of the variation can be explained by the first 2 principal components. The first component is made up of a weighted difference between (x1,x3) and (x2,x4,x5). The second component is also a weighted difference between (x1,x2,x3,5) and (x4)

I would recommend using the PCA.2s because it explains a lot more variation with fewer variables.

\newpage

3)

```{r q3, warning=FALSE, comment=NA, message=FALSE}
## a)
X3 = read.table("T1-10.DAT", quote="\"", comment.char="")
colnames(X3) = c("Breed", "SalePr", "YrHgt", "FtFrBody", "PrctFFB", "Frame", 
                 "BkFat", "SaleHt", "SaleWt")

head(X3)

## Sample variance
S3 = var(X3[, 2:9])

## pricipal components
## the first 3 PC explain 86% of the variation
pca.3r = prcomp(X3[, 2:9], center = TRUE, scale. = TRUE)
summary(pca.3r)

## b) The scree plot significantly levels off after 2 components however
##    the first 2 components only account for 74% of the data so I would
##    include the 3rd component since it will get you to 86%
plot(x = 1:8, y = eigen(S3)$values, type = "b", main = "Scree Plot")

## c)
pca.3r

## PC1 could be viewed as a weighted difference between all variables and 
##     BkFat, or BkFat could small enough that we can drop it and say we 
##     have a weighted average between all variables
##
## PC2 is a weighted difference between (SalePrc, BkFat, SaleWt) and (PrctFB) 
##     after we drop the variables near 0

## d)
PC.3 = as.matrix(X3[, 2:9]) %*% pca.3r$rotation
dta = data.frame(BREED = X3$Breed, PC.3[, 1:2])

plot(x = dta$PC1, y = dta$PC2, col = factor(dta$BREED))
legend(x = -1800, y = -1000, unique(dta$BREED), col = 1:3, pch = 1)

```

\newpage

```{r q4, warning=FALSE, comment=NA, message=FALSE}

## You can definetely see that breed 5 is difference from breed 8. And Breed 1
## has a lot of variation. The 3 outliers are in terms of price.  These are
## the top 3 priced bulls. The summary statistics indicate that they are in the
## upper percentiles (except for the fat measurements) compared to their same 
## breed.

## outliers 
outliers = order(dta$PC1)[1:3]
X3[outliers, ]


## Breed 1
summary(subset(X3, Breed == 1, 2:9))
        
## Breed 8
summary(subset(X3, Breed == 8, 2:9))

## Breed 5
summary(subset(X3, Breed == 5, 2:9))


```





















