---
output:
  pdf_document:
    fig_height: 5
    fig_width: 8
    highlight: pygments
    latex_engine: xelatex
mainfont: Calibri
sansfont: Calibri
fontsize: 12pt
geometry: margin=1in
---

**Homework 04**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 636-720**  

\newpage

1)

  a) All 3 variables appear to be normally distributed based on how the points follow close to the normal probability line. The shapiro-wilks test also supports the claim that all 3 variables are normally distributed

```{r a, comment=NA, fig.height=3}

sweat = read.table("T5-1.DAT", quote="\"", comment.char="")
colnames(sweat) = c("SR", "NA", "K")
head(sweat)
par(mfrow = c(1, 3))
qqnorm(sweat$SR); qqline(sweat$SR)
qqnorm(sweat$`NA`); qqline(sweat$`NA`)
qqnorm(sweat$K); qqline(sweat$K)

shapiro.test(sweat$SR)
shapiro.test(sweat$`NA`)
shapiro.test(sweat$K)

```

\newpage

b)

```{r b, comment=NA}

## Covariance Matrix
(CV = cov(sweat))

## confidence region
(C2 = qchisq(.95, 3))

## mu center
colMeans(sweat)

## axis
(eigen.vect = eigen(CV)$vectors)

## half lengths
eigen.vals = eigen(CV)$values
C2/sqrt(eigen.vals)


```

\newpage

c) Simultaneous Confidence Interval

```{r c, comment=NA}

## Confidence Region
(C2 = (20 - 1) * 3 * qf(.95, 3, 17) / (20 - 3))

## Simultaneous
(ci_sim_sr = colMeans(sweat)[1] + c(-1, 1) * sqrt(C2 * CV[1, 1] / 20))
(ci_sim_na = colMeans(sweat)[2] + c(-1, 1) * sqrt(C2 * CV[2, 2] / 20))
(ci_sim_k = colMeans(sweat)[3] + c(-1, 1) * sqrt(C2 * CV[3, 3] / 20))

```

d) Bonferroni Confidence Interval
  
```{r d, comment=NA}

## Bonferroni
(ci_bon_sr = colMeans(sweat)[1] + c(1, -1) * qt(.95 / 6, 19) * sqrt(CV[1, 1] / 20))
(ci_bon_na = colMeans(sweat)[2] + c(1, -1) * qt(.95 / 6, 19) * sqrt(CV[2, 2] / 20))
(ci_bon_k = colMeans(sweat)[3] + c(1, -1) * qt(.95 / 6, 19) * sqrt(CV[3, 3] / 20))

```

\newpage

e) Our p-value is .3 which is much greater than the .05 cuttoff so we do not reject the null and conclude that $\mu_0 = [4, 45, 10]$ could be the true mean.

```{r e, comment=NA, message=FALSE, warning=FALSE}
library(Hotelling)
library(DescTools)

## hypothesis test
mu_0 = c(4, 45, 10)

## test statistic
T2 = 20 * t(colMeans(sweat) - mu_0) %*% solve(CV) %*% (colMeans(sweat) - mu_0)
((20 - 3) / ((20 - 1) * 3)) * T2

## critical value
qf(.05, 3, 17)

## pvalue
1 - pf((20 - 3) * T2 / ((20 - 1) * 3), 3, 17)

HotellingsT2Test(x = sweat, mu = mu_0, test = 'f')


```

\newpage

f) Since our test statistic is less than the perimeter distance of our 95% confidence ellipsoid we know that $\mu_0 = [4, 45, 10]$ is inside the region.

```{r f, comment=NA}

T2 < C2

```


g)

```{r g, comment=NA}

## Center the sample around the null hypothesis
X_0 = sweat - rep(1, 20) %*% t(colMeans(sweat)) + rep(1, 20) %*% t(mu_0)

## results
A = rep(NA, 500)

## set the seed
set.seed(101)
samps = sample(1:20, size = 500*20, replace = TRUE)
samps = matrix(samps, nrow = 500, byrow = TRUE)

for(i in 1:500) {
  ## Generate bootstrap sample under H_0.
  X_b = X_0[samps[i, ], ]
  
  ## Compute the variance of the sample
  S_b = cov(X_b)
  
  ## Calculate the variance
  S_b = (1/19) * t(as.matrix(X_b - colMeans(X_b))) %*% (as.matrix(X_b - colMeans(X_b)))
  S_0 = (1/19) * t(as.matrix(X_b - mu_0)) %*% (as.matrix(X_b - mu_0))
    
  ## Compute test statistic
  A[i] = ( det(S_b) / det(S_0) )^10
}

## pvalue
length(which(A > 1) == TRUE) / 500

```





