---
output:
  pdf_document:
    fig_height: 5
    fig_width: 8
    highlight: pygments
    latex_engine: xelatex
mainfont: Calibri Light
sansfont: Calibri Light
fontsize: 12pt
geometry: margin=1in
---

**Homework 01**  
**Joseph Blubaugh**  
jblubau1@tamu.edu  
**STAT 636-720**  

\newpage


1)

```{r q1, comment=NA}

oxygen = 
  read.table("C:/Users/Joseph/Projects/learning/Statistics/STAT_636/Data/textbook/T6-12.DAT", 
             quote="\"", 
             comment.char="")
colnames(oxygen) = c("rest_L.min", "rest_mL.kg.min", "strenuous_L.min", 
                     "strenuous_mL.kg.min", "gender")

## a)
## Sample Averages
apply(X = oxygen[, 1:4], MARGIN = 2, FUN = mean)

## Sample Standard Deviations
apply(X = oxygen[, 1:4], MARGIN = 2, FUN = sd)

## b)
pairs(oxygen)

## c)
coplot(rest_L.min ~ strenuous_L.min | gender, data = oxygen)

```

a) The resting oxygen measurement means and standard deviation are lower than that measurements for the strenuous activity. Strenuous mL/kg/min is much higher and has more variation than the Resting mL/kg/min measurement.

b) Resting L/min appears to be positively correlated with both Rest mL/kg/min and Strenuous L/min. Strenuous L/min also appears to be positively correlated with Strenuous mL/kg/min The observations of the female with a resting mL/kg/min of 11.05 appears to be an outlier and shows up as such on several plots.

c) Visually there appears to be no obvious relationship between Rest L/min and Strenuous L/min for females, but the relationship appears to be positively correlated for men.


2)
  a)
  
\begin{displaymath}
x = 
\begin{bmatrix}
    x_1 \\
    x_2 \\
\end{bmatrix}
\mu = 
\begin{bmatrix}
    \mu_1 \\
    \mu_2 \\
\end{bmatrix}
\Sigma = 
\begin{bmatrix}
    \sigma_{11} & \sigma_{12} \\
    \sigma_{12} & \sigma_{22} \\
\end{bmatrix}
\end{displaymath}

\begin{displaymath}
d(x, \mu) = \sqrt{(x - \mu)' \Sigma^{-1} (x - \mu)} \\
\end{displaymath}

\begin{displaymath}
d(x,\mu) = \sqrt{
\begin{bmatrix}
    x_1 - \mu_1 & x_2 - \mu_2 \\
\end{bmatrix}
\begin{bmatrix}
    \sigma_{11} & \sigma_{12} \\
    \sigma_{12} & \sigma_{22} \\
\end{bmatrix}^{-1}
\begin{bmatrix}
    x_1 - \mu_1 \\
    x_2 - \mu_2 \\
\end{bmatrix}}
\end{displaymath}

\begin{displaymath}
 = \sqrt{\sigma_{11}(x_1 - \mu_1)^2 + 2 \sigma_{12}(x_1 - \mu_1)(x_2 - \mu_2) + \sigma_{22}(x_2 - \mu_2)^2} \\
\end{displaymath}

\begin{displaymath}
a_{11} = \sigma_{11}, a_{12} = \sigma_{12}, a_{22} = \sigma{22}
\end{displaymath}

b i.

```{r q2, comment=NA}

mu = c(1, -1)
Sigma = matrix(c(1, -1.6, -1.6, 4), nrow = 2)

mvn = function(x) {
  constant = 1/((2 * pi)^(2/2) * sqrt(det(Sigma)))
  scalar = exp(-t(x - mu) %*% solve(Sigma) %*% (x - mu)/2)
  
  out = constant * scalar
  
  return(as.numeric(out))
}


x1 = seq(-2, 4, length = 75)
x2 = seq(-13, 11, length = 75)

results = matrix(NA, nrow = 75, ncol = 75)

for (i in 1:75) {
  for (j in 1:75) {
    results[i, j] = mvn(c(x1[i], x2[j]))
  }
}

persp(results, ticktype = "detailed")


```

\newpage

ii.

\begin{displaymath}
\mu = 
\begin{bmatrix}
    1 \\
    -1 \\
\end{bmatrix}
\Sigma = 
\begin{bmatrix}
    1.0 & -1.6 \\
    -1.6 & 4 \\
\end{bmatrix}
\end{displaymath}

\begin{displaymath}
d(P,Q) = \sqrt{\sigma_{11}(x_1 - \mu_1)^2 + \sigma_{12}(x_1 - \mu_1)(x_2 - \mu_2) + \sigma_{22}(x_2 - \mu_2)^2}
\end{displaymath}

\begin{displaymath}
O = (0, 0),  P = (0, -2)
\end{displaymath}

**Statistical Distance**

\begin{align*}
d(O,Q) =& \sqrt{1(0 - 1)^2 - 1.6(0 - 1)(0 + 1) + 4(0 + 1)^2)} \\
       =& \sqrt{1 + 1.6 + 4} \\
       =& \sqrt{6.6} = 2.569 \\
\\
d(P,Q) =& \sqrt{1(0 - 1)^2 - 1.6(0 - 1)(-2 + 1) + 4(-2 + 1)^2)} \\
       =& \sqrt{1 - 1.6 + 4} \\
       =& \sqrt{3.4} = 1.843 \\
\end{align*}

**Straightline Distance**

\begin{align*}
d(O,Q) =& \sqrt{(0 - 1)^2 + (0 + 1)^2} \\
       =& \sqrt(2) \\
       =& 1.41 \\
\\
d(P,Q) =& \sqrt{0 - 1)^2 + (-2 + 1)^2} \\
       =& \sqrt{2} \\
       =& 1.41
\end{align*}

**Conclusion**: Using the straightline distance both P and O are equally distant from Q, but when taking into accoun the variances, P is actually closer to Q than O is.

\newpage

iii.  

For $R_P$, $x_2$ appears to no longer by centered on $P$ and as such will have less volume under the curve than $R_O$ which is centered on $O$ for both $x_1$ and $x_2$, therefore $P(x \in R_O) > P(x \in R_P)$.



























