---
title: ''
output:
  pdf_document: default
---
  
Joseph Blubaugh  
STAT 630  
HW10  
  
  
##### 7.1.3  
...
```{r q1}
location_normal = 1 - pnorm(0, 0, sqrt(1))
prior = 1 - pnorm(0, 1, sqrt(10))
location_normal * prior

```

##### 7.1.4  
...
\begin{align*}
  Poisson(\lambda) &= \frac{\lambda^x}{x!} e^{-\lambda} \\
  Gamma(a,b) &= \frac{b^a x^{a - 1}}{\gamma(a)} e^{-b x} \\
  \\
  f(x, \theta) &= Poisson(\lambda) * Gamma(a,b) \\
  &= \frac{b^a x^{a - 1}}{\gamma(a)} e^{-b x} \frac{\lambda^x}{x!} e^{-\lambda} \\
  \\
  m(x) &=  \frac{b^a x^{a - 1}}{\gamma(a)} \frac{1}{x! e^x} e^{-b x} \int_0^1 \frac{x! e^x}{\lambda^{x-1}} \frac{\lambda^x}{x!} e^{-\lambda} d\lambda   \\
  \\
  Posterior &= f(x, \lambda) \\
  &= \frac{\frac{b^a x^{a-1}}{\gamma(a)} e^{-b x} \frac{\lambda^x e^{-\lambda}}{x!}}{\frac{b^a x^{a-1}}{\gamma(a)} e^{-b x} \frac{\lambda^{x-1}}{x! e^x}} \\
  &= \frac{x! \lambda^x e^{-\lambda x}}{\lambda^{x-1} x!} \\
  &= \lambda e^{-\lambda x} \\
\end{align*}

(b) Find the posterior mean, posterior mode, and posterior variance.  
PosteriorMean = $1/\lambda$  
PosteriorMode = 0  
PosteriorVariance = $1/\lambda^2$  
  
  
##### 7.1.9  
...
\begin{align*}
  Uniform[.4, .6] &= 1 / (.6 - .4) \\
  Binomial(n, \theta) &= {n \choose x} \theta^x (1 - \theta)^{n - x} \\
  f(x,\theta) &= Binomial(n, \theta) * Uniform[.4, .6] \\
  \\
  m(x) &= \frac{1}{.6 - .4} (n + 1) \theta^n \int_0^1 \frac{{n \choose x} \theta^n (1 - \theta)^{n - x}}{(n + 1) \theta^n} \\
  \\
  posterior &= \frac{(n + 1) \theta^n I[.4, .6] \theta}{(.6^{n + 1} - .4^{n - 1})} \\
\end{align*}  

##### 7.1.14  
...  
```{r q3, fig.height=3, fig.width=5}
mu = function(x, gam, var, n, mu0) ((1 / gam) + (n / var))^-1 * ((mu0/gam) + (n/var)) * x
sd0 = function(n, var, gam) ((1 / gam) + (n / var))^-1

x = rnorm(n = 1000, 
          mean = mu(x = 8.2, gam = 1, var = 1, n = 10, mu0 = 2), 
          sd = sd0(n = 10, var = 1, gam = 1))

error = sd(x) / sqrt(length(x))
mean(x) - error; mean(x) + error

hist(x)
```


##### 7.2.1  
...
\begin{align*}
  PosteriorMean &= \frac{\theta^{(n \bar{x} + \alpha - 1)} (1 - \theta)^{n (1 - \bar{x}) + \beta - 1}}{n} \\
  \\
  n = 40, \bar{x} &= .25, \alpha = 1, \beta = 1 \\
  \\
  Posterior Mean &= \frac{\theta^{(40 * .25 + 1 - 1)} (1 - \theta)^{40 (1 - .25) + 1 - 1}}{40} \\
  &= \frac{\theta^{10} (1 - \theta)^{30}}{40} \\
\end{align*}

##### 7.2.2  
since the distribution is symmetrical the mode and the mean are equal
\begin{align*}
  \psi &= \mu + \sigma_0 z_.75 \\
  &= \mu + 1 (.67) \\
  \\
  \psi &= ( \frac{1}{\gamma_0^2} + \frac{n}{\sigma_0^2})^{-1} ) ( \frac{\mu_0}{\gamma_0^2} + \frac{n}{\sigma_0^2} ) + .67 \\
  \\
  \gamma_0^2 = 2, \sigma_0^2 &= 1, \mu_0 = 0, n = 10 \\
  \\
  \psi &= ( \frac{1}{2} + \frac{n}{1})^{-1} ) ( \frac{0}{2} + \frac{10}{1} ) + .67 \\
  &= .67 \\
\end{align*}


##### 7.2.10
\begin{align*}
  X &= Exponential(\lambda) = \lambda e^{-\lambda x} \\
  \lambda &= Gamma(\alpha, \beta) = \frac{\beta^{\alpha} x^{\alpha - 1}}{\gamma(\alpha)} e^{-\beta x} \\
  \\
  f(x | \lambda) &= Exponential(\lambda) * Gamma(\alpha, \beta) \\
  &= \frac{\beta^{\lambda} x^{\alpha - 1}}{\gamma(\alpha)} \lambda e^{\lambda \beta x^2} \\
  \\
  m(x) &= \frac{\beta^{\alpha} x^{\alpha - 1}}{\gamma(\alpha)} e^{-\beta x} \int_0^1 \frac{1}{e^{-\beta x}} \lambda e^{\lambda \beta x^2}, d\lambda \\
  \\
  Posterior &= \lambda e^{-\lambda \beta^2 x^3} \\
  Expectation &= \int_0^1 \lambda^2 e^{-\lambda \beta^2 x^3} \\
  &= -( \frac{1}{\beta^2 x^3} + \frac{2}{\beta^4 x^6} + \frac{2}{\beta^6 x^9}) e^{-\beta x^3} \\
\end{align*}

##### 6.3.27
$1 - 2 \phi ( \frac{\bar{x} - \mu}{\sigma - \sqrt{n}})$

##### 8.2.5
a) 0
b) $1 - (1/\theta)$  

##### Additional A
\begin{align*}
  binomial(100, \theta) &= {100 \choose x} \theta^x (1 - \theta)^(n - x) \\
  \\
  \beta(\mu') &= \frac{.1 - \mu}{.02 / \sqrt{100}} \le -Z_{1 - \alpha} \\
  \mu &= .1 \\
  \alpha &= 0 \\
\end{align*}
```{r qa, fig.height=3, fig.width=5}
theta = 40:50
bin = choose(100, 50) * theta^50 * (1 - theta)^(100 - 50)
abs.bin = abs(bin - 50)
plot(bin, type = "l")

```


##### Additional B
a) best order x = {4, 1, 2, 3}  
 

|    x        |  1  |  2  |  3  |  4  |  
|-------------|-----|-----|-----|-----|  
| f0(x)       | 0.2 | 0.3 | 0.3 | 0.2 |  
| f1(x)       | 0.1 | 0.4 | 0.1 | 0.4 |  
| f1(x)/f0(x) | .5  | 1.3 | 0.3 | 2.0 |  
|-------------|-----|-----|-----|-----|
 
 
b) 
  at a = .2 -> x = 1,2,3,4 
  at a = .5 -> x = 1,2,4

c) 
  at a = .2 -> power = .2 + .3 + .3 + .2 = 1
  at a = .5 -> power = .2 + .3 + .2 = .8

d) 
  outcomes x = 1 and x = 3





















